Comparison of Catalogs

### 2024-07-21 compare/compare.qmd
try sigma

[^1]: Fitting can be done using the uncertainties as weights. To get the standard weighting of 1/unc^2, assuming Gaussian errors, the weights to pass to the fitting are 1/unc.
[^2]: In simple linear regression, $R^2$ is the square of the Pearson correlation coefficient $\rho$.

```{python}
def compare_data(x, y, unc=None):
    """
    Performs a linear comparison between two datasets.

@@

        # fit the data with the fitter
        fitted_line = fit(line_init, x_data, y_data, weights=w)
        slope = fitted_line.slope.value
        intercept = fitted_line.intercept.value

@@

        return slope, intercept, r2, corr
    except Exception:
        return 0, 0, 0, 0
```

@@

```{python}
def scatter_plot(x, y, xerr = None, yerr = None):
    """
    Generates a scatter plot of two datasets with optional error bars, fits a linear model to the data, 
    and displays the fitted line on the plot.

@@

    # Perform linear comparison between the datasets
    if yerr is not None:
        unc = yerr
    elif yerr is None and xerr is not None:
        unc = xerr
    else:
        unc = None
    c, m, r2, corr = compare_data(x, y, unc)

    # Plot the data with error bars
    # errorbar -> needs np.ma -> no units
    plt.errorbar(x_data, y_data, xerr = xerr_d, yerr = yerr_d, alpha = 0.3, c = "blue", label = "Data", fmt=".")

    # Plot the fitted line with unit support
    if c!=0:
        plt.plot(x, c*x.value+ m, c = "red", label = f"Fit: {y.info.name}$_H $= {c:.2f}$\cdot${x.info.name}$_L${m:+.2f}\n $R^2=$ {r2:.2f}")

@@

ttype_corr = round(compare_data(dt["TType"], dt["T"])[3], 4)
inc_corr = round(compare_data(dt["inc"], dt["INCL"])[3], 3)
a26_corr = round(compare_data(dt["a26_1"], dt["A26_1a"])[3], 3)
```

|      LCV      |        HECATE        |                        Description                        | Pearson Correlation \[-1,1\] |
|:---------------:|:---------------:|:-------------------:|:---------------:|
|     TType     |   T (with errors)    | Numerical Hubble type following the de Vaucouleurs system |    `{python} ttype_corr`     |
|      inc      |          INCL           |                     Inclination (deg)                     |     `{python} inc_corr`      |
| a26_1 (Major) | A26_1a (Semi-major axis) |                 angular diameter (arcmin)                 |              `{python} a26_corr`               |

@@

scatter_plot(dt["TType"], dt["T"], yerr = dt["E_T"])
scatter_plot(dt["inc"], dt["INCL"], yerr = dt["E_V"])
scatter_plot(dt["a26_1"], dt["A26_1a"], yerr = dt["E_V"])

@@

```{python}
#| echo: false
logKLum_corr = round(compare_data(dt["logKLum"], dt["logL_K"])[3], 3)
```
|   LCV   | HECATE | Description | Pearson Correlation \[-1,1\] |
|:-------:|:------:|:-----------:|:----------------------------:|
| logKLum | logL_K |             |      `{python} logKLum_corr` |

@@

|         LCV         | HECATE | Description |
|:-------------------:|:------:|:-----------:|
| mag_B (with errors) |  BT (with errors)      |             |
|       Kmag   |     K   | 2MASS band magnitude  (both)          |
|                     |        |             |
## Absorbsion

```{python}
#| echo: false
ab_corr = round(compare_data(dt["AB"], dt["AG"])[3], 3)
#ab_int_corr = round(compare_data(dt["AB_int"], dt["AI"])[3], 3)
```
|  LCV   | HECATE |                  Description                   | Pearson Correlation \[-1,1\] |
|:------:|:------:|:----------------------------------------------:|:----------------------------:|
|   AB   |   AG   |    Galactic extinction/absorption in B band    |      `{python} ab_corr`      |
| AB_int |   AI   | Internal/Intrisic B band extinction/absorption |   0   |
```{python}
scatter_plot(dt["AB"], dt["AG"])#, scatter_plot(dt["AB_int"], dt["AI"])
```
## SFR

>>>>

[^1]: Fitting can be done using the uncertainties as weights. To get the standard weighting of 1/unc\^2, assuming Gaussian errors, the weights to pass to the fitting are 1/unc.

[^2]: In simple linear regression, $R^2$ is the square of the Pearson correlation coefficient $\rho$.

```{python}
def compare_data(x, y, unc=None, sigma = False):
    """
    Performs a linear comparison between two datasets.

@@

        # fit the data with the fitter
        # check if sigma
        if sigma is False:
            fitted_line = fit(line_init, x_data, y_data, weights=w)
            outliers = None
        else:
            or_fit = fitting.FittingWithOutlierRemoval(fit, sigma_clip, niter=3, sigma=3)
            fitted_line, outliers = or_fit(line_init, x, y, weights=w)
        slope = fitted_line.slope.value
        intercept = fitted_line.intercept.value

@@

        return slope, intercept, r2, corr, outliers
    except Exception:
        return 0, 0, 0, 0, None
```

@@

```{python}
def scatter_plot(x, y, xerr = None, yerr = None, sigma = False):
    """
    Generates a scatter plot of two datasets with optional error bars, fits a linear model to the data, 
    and displays the fitted line on the plot.

@@

    # Perform linear comparison between the datasets
    if yerr is not None:
        unc = yerr
    elif yerr is None and xerr is not None:
        unc = xerr
    else:
        unc = None
    c, m, r2, corr, outliers = compare_data(x, y, unc, sigma)

    # Plot the data with error bars
    # errorbar -> needs np.ma -> no units
    if outliers is not None:
        filtered_data = np.ma.masked_array(y_data, mask = outlier)
        plt.scatter(x, filtered_data,  label="Fitted Data", c = "blue", fillstyle = "filled")
        plt.errorbar(x, y, yerr=yunc, fillstyle="filled", c = "black", label="Clipped Data", fmt = "o")
    else:
        plt.errorbar(x_data, y_data, xerr = xerr_d, yerr = yerr_d, alpha = 0.3, c = "blue", label = "Data", fmt = ".")

    # Plot the fitted line 
    if c!=0:
        plt.plot(x, c*x.value+ m, c = "red", label = f"Fit: {y.info.name}$_H $= {c:.2f}$\cdot${x.info.name}$_L${m:+.2f}\n $R^2=$ {r2:.2f}")

@@

ttype_corr = round(compare_data(dt["TType"], dt["T"])[3], 4)
inc_corr = round(compare_data(dt["inc"], dt["INCL"], sigma = True)[3], 3)
a26_corr = round(compare_data(dt["a26_1"], dt["A26_1a"], sigma =True)[3], 3)
```

|      LCV      |          HECATE          |                        Description                        | Pearson Correlation \[-1,1\] |
|:---------------:|:---------------:|:------------------:|:---------------:|
|     TType     |     T (with errors)      | Numerical Hubble type following the de Vaucouleurs system |    `{python} ttype_corr`     |
|      inc      |           INCL           |                     Inclination (deg)                     |     `{python} inc_corr`      |
| a26_1 (Major) | A26_1a (Semi-major axis) |                 angular diameter (arcmin)                 |     `{python} a26_corr`      |

@@

scatter_plot(dt["TType"], dt["T"], yerr = dt["E_T"])
scatter_plot(dt["inc"], dt["INCL"], yerr = dt["E_V"], sigma =True)
scatter_plot(dt["a26_1"], dt["A26_1a"], yerr = dt["E_V"])

@@

```{python}
#| echo: false
logKLum_corr = round(compare_data(dt["logKLum"], dt["logL_K"])[3], 3)
```

|   LCV   | HECATE | Description | Pearson Correlation \[-1,1\] |
|:-------:|:------:|:-----------:|:----------------------------:|
| logKLum | logL_K |             |   `{python} logKLum_corr`    |

@@

```{python}
#| echo: false
mag_B_corr = round(compare_data(dt["mag_B"], dt["BT"],unc = dt["E_BT"], sigma = True)[3], 3)
Kmag_corr = round(compare_data(dt["Kmag"], dt["K"], unc = dt["E_K"])[3], 3)
```

|         LCV         |      HECATE      |         Description         | Pearson Correlation \[-1,1\] |
|:---------------:|:-------------:|:-------------:|:------------------------:|
| mag_B (with errors) | BT (with errors) |                             |    `{python} mag_B_corr`     |
|        Kmag         |        K         | 2MASS band magnitude (both) |     `{python} Kmag_corr`     |

```{python}
scatter_plot(dt["mag_B"], dt["BT"], yerr = dt["E_BT"], sigma = True)
scatter_plot(dt["Kmag"], dt["K"], yerr = dt["E_K"])
```

## Absorbsion

```{python}
#| echo: false
ab_corr = round(compare_data(dt["AB"], dt["AG"])[3], 3)
#ab_int_corr = round(compare_data(dt["AB_int"], dt["AI"])[3], 3)
```

|  LCV   | HECATE |                  Description                   | Pearson Correlation \[-1,1\] |
|:-----------:|:-----------:|:----------------------------:|:---------------:|
|   AB   |   AG   |    Galactic extinction/absorption in B band    |      `{python} ab_corr`      |
| AB_int |   AI   | Internal/Intrisic B band extinction/absorption |              0               |

```{python}
scatter_plot(dt["AB"], dt["AG"])#, scatter_plot(dt["AB_int"], dt["AI"])
```

## SFR
